IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/real.tex
Pass over parts of real.tex

In the introduction, we stated that the original motivation behind
developing this embedding was in order to formally specify a
coarse-grained information flow control language for JavaScript.
%
Indeed, this work serves as the theoretical foundation for the
JavaScript IFC systems built atop Node.js and Gecko layout engine.
%
In this section, we consider the application of our embedding to
JavaScript for these two platforms, and two other real world target
languages, Haskell, and C.
%
Our goal is to show the flexibility of the embedding in settings with
vastly different features and properties, as well as discuss the
semantic gap that must be overcome in some cases to apply our formalism
to other systems.
%
%% Two of the systems we describe have been implemented: the Haskell
%% system~\cite{lio} and the SWAPI~\cite{swapi} system; we leave the
%% implementation of the IFC system for C to future work.

@@

JavaScript, as specified by
ECMAScript~\cite{ecma}, does not have any built-in
functionality for I/O
(we denote this language as |targetLangJS|).
%
When JavaScript is embedded in web browser engines, such as Google's
Blink or Mozilla's Gecko, or in server-side platforms such as Node.js,
additional APIs such as the Document Object Model (DOM) or the file
system get exposed as part of the runtime system.
%
In a similar manner, we can directly implement the embedded language
|specLangJS roundrobinf| by exposing (only) our IFC primitives as part
of the runtime and running multiple instances of the JavaScript VM in
separate OS-level threads.\footnote{
  Indeed, the Gecko implementation of this work does this for Web
  Workers that run operate OS threads~\cite{cowl}.
}
%
Unfortunately, this becomes very costly when an application relies on
many tasks.
%
(Consider, for instance, a server-side web application that wishes to
create a new task for each HTTP request.)

However, this issue of performance and JavaScript isolation is not
unique to our work. 
%
In particular, layout engines rely on the isolation of content from
different origins (e.g., two iframes of different origins) to
implement the \emph{same-origin policy}.
%
Since creating an OS thread for each iframe is expensive, both
Google's V8 and Mozilla's SpiderMonkey provide a means for running
JavaScript code in isolation, on disjoint sub-heaps of the same OS
thread.
%
In V8, this unit of isolation is called a \emph{context}~\tocite{}; in
SpiderMonkey, it called a \emph{compartment}~\tocite{}.
%
Each context/compartment is associated with a global object, which, by
default, implements the JavaScript standard library (e.g.,
\verb|Object|, \verb|Array|, etc.).
%
(Gecko and Node.js expose features such as the DOM and
file system by extending this global object.)
%
Naturally, our notion of a task for JavaScript directly corresponds to
a context/compartment.
%
Below we describe some of the implementation details specific to
Node.js and Gecko.

@@

We implemented |specLangJS seqf| for Node.js in the form of a module
(library), which does not modify the Node.js or V8 code-base in any
way.
%
Our implementation provides a library for creating new tasks, i.e.,
contexts whose global object only contains the standard JavaScript
library and our IFC primitives (|send|, |sandbox|, etc.).
%
In turn, these tasks can create new tasks with |sandbox|.
%
In our formal treatement, |sandbox| is defined with |klone(tS) = tS0|,
where |tS0| is the global object corresponding to the standard
JavaScript library and our IFC primitives.

\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{figs/node}}
\caption{\label{fig:node} Retrofitting Node.js JavaScript with IFC.
This example shows how our trusted monitor (left) is used to mediate
communication between two tasks for which IFC is enforced (right).}
\end{figure}
%
As shown in Figure~\ref{fig:node}, each IFC API call from a task is mediated
by the trusted library code (the main Node.js context), that tracks
the current label, messages, etc. of each task.
%
(This task is part of the trusted computing base (TCB) and has access
to the full Node.js API.)
%
While V8 provides a means for sharing objects between contexts (used
by same-origin contexts in Blink), our system conservatively restricts
the kinds of messages that can be exchanged by tasks.
%
Concretely, and as mentioned in Section~\ref{sec:concrete}, we
restrict |send| (and |sandbox|) to sending string values.
%
In our formalization, this amounts to restricting the IFC language rule
for |send| such that only strings can be shared:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  iv) , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' iv)), ldots)
.->
iS'; sched step (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
%
Our implementation, however, is accompanied by a library that provides
a more flexible standard library, which, among other things,
automatically marshalls JSON objects to/from strings on each
|send|/|recv|.

@@

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary syscalls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
For instance, this isolation can be achieved using Dune's hardware protection
mechanisms~\cite{Belay:2012:DSU:2387880.2387913}, similar to a simpler
version of Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611} with an information flow control
policy.
%
Using page tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates--tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept syscalls performed by
a task and only permit those corresponding to our IFC language (such as
|getLabel| or |send|).
%
Dunes hardware protection mechanism allows us to provide a concrete
implementation that is efficient and relatively simple to reason
about, but other sandboxing mechanisms could be used in place of Dune.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a syscall.
%
Creating a new task with the |sandbox| syscall corresponds to
\emph{forking} a process.  Here, page tables limit what memory
a task can access: we can ensure there will be no memory (effectively
defining |klone(tS)
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process.)
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these syscalls are
meaningless.

>>>>

In the introduction, we stated that the original motivation behind
this work was to formally specify a
coarse-grained information flow control language for JavaScript.
%
Indeed, our formalism serves as the theoretical foundation for two
IFC systems for JavaScript, built on top of Node.js as well as
the web browser engine Gecko.

In this section, we describe these two implementations.  Furthermore,
we consider how our formalism could be applied to the C programming
language, and connect it to a previous IFC system for Haskell.
%
Our goal is to show the flexibility of the embedding in settings with
vastly different properties, as well as to discuss the
semantic gap that must be overcome in some cases to apply our formalism
to other systems.
%
%% Two of the systems we describe have been implemented: the Haskell
%% system~\cite{lio} and the SWAPI~\cite{swapi} system; we leave the
%% implementation of the IFC system for C to future work.

@@

JavaScript, as specified by
ECMAScript~\cite{ecma}, does not have any built-in
functionality for I/O.
%(we denote this language as |targetLangJS|).
%
For this language, which we denote by |targetLangJS|, the IFC system
|specLangJS roundrobinf| can be implemented by exposing IFC primitives
to JavaScript as part of the runtime, and running multiple instances
of the JavaScript virtual machine in separate OS-level threads.
%
Unfortunately, this becomes very costly when an application relies on
many tasks.
%
For instance, consider a server-side web application that
creates a new task for each HTTP request.  The overhead would
quickly become impractical.

However, this issue of performance and JavaScript isolation is not
unique to our work, and has been solved by browser engines,
which
rely on the isolation of content from
different origins (e.g., two iframes of different origins) to
implement the same-origin policy.
%
Since creating an OS thread for each iframe is expensive, both
Google's V8 and Mozilla's SpiderMonkey provide a means for running
JavaScript code in isolation in a single OS thread,
on disjoint sub-heaps.
%
In V8, this unit of isolation is called a \emph{context}~\tocite{}; in
SpiderMonkey, it is called a \emph{compartment}~\tocite{}.
%
Each context/compartment is associated with a global object, which, by
default, implements the JavaScript standard library (e.g.,
\verb|Object|, \verb|Array|, etc.).
%
Naturally, our notion of a task for JavaScript directly corresponds to
a context/compartment.

When JavaScript is embedded in web browser engines, such as Google's
Blink or Mozilla's Gecko, or in server-side platforms such as Node.js,
additional APIs such as the Document Object Model (DOM) or the file
system get exposed as part of the runtime system.
These features are exposed by extending the global object, just like
for the standard library.  For this reason, it is easy to modify
these systems to remove arbitrary external effects when implementing
an IFC system, and reintroducing important effects in a safe manner.

@@

We have implemented |specLangJS seqf| for Node.js in the form of a module
(library), which does not modify the Node.js or V8 code-base in any
way.
%
Our implementation provides a library for creating new tasks, i.e.,
contexts whose global object only contains the standard JavaScript
library as well as our IFC primitives (such as |send| and |sandbox|).
%
In turn, these tasks can create new tasks with |sandbox|.
%
In our formal treatement, |sandbox| is defined with |ap klone tS = tS0|,
where |tS0| is the global object corresponding to the standard
JavaScript library and our IFC primitives.

\begin{figure}
\centerline{\includegraphics[width=\columnwidth]{figs/node}}
\caption{\label{fig:node} Retrofitting Node.js JavaScript with IFC.
This example shows how our trusted monitor (left) is used to mediate
communication between two tasks for which IFC is enforced (right).}
\end{figure}
%
All IFC primitives are mediated by the trusted library code (executing
as the main Node.js context), which tracks the current label, messages,
etc.\ of each task.  An example for |send|/|recv| is shown in
Figure~\ref{fig:node}.

%While V8 provides a means for sharing any object between contexts (used
%by same-origin contexts in Blink),
Our system conservatively restricts
the kinds of messages that can be exchanged by tasks by
only allowing string values to be exchanged via |send| (and |sandbox|).
%
In our formalization, this amounts to restricting the IFC language rule
for |send| in the following way:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  iv) , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' iv)), ldots)
.->
iS'; sched step (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
%
For convenience, we have implemented a library that provides
more flexible primitives, which, among other things,
automatically marshalls JSON objects to/from strings on each
|send|/|recv|.

@@

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary system calls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
For instance, this isolation can be achieved using Dune's hardware protection
mechanisms~\cite{Belay:2012:DSU:2387880.2387913}, similar to a simpler
version of Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611} with an information flow control
policy.
%
Using page tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates--tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept system
calls performed by
a task and only permit those corresponding to our IFC language (such as
|getLabel| or |send|).
%
Dunes hardware protection mechanism allows us to provide a concrete
implementation that is efficient and relatively simple to reason
about, but other sandboxing mechanisms could be used in place of Dune.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a system call.
%
Creating a new task with the |sandbox| system call corresponds to
\emph{forking} a process.  Using page tables we can ensure that
there will be no shared memory
(effectively
defining |ap klone tS
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process.)
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these system calls are
meaningless.

We leave an actual implementation of an IFC system for C as
future work.
